\section{Object Oriented Programming}

\begin{frame}[plain]
\sectionpage
\end{frame}



\begin{frame}{Introduction}
Python is an object-oriented programming language, and classes are fundamental to creating and managing objects. A class defines the blueprint for an object, encapsulating data (attributes) and behavior (methods).
\end{frame}



\begin{frame}[fragile]
\frametitle{Defining a Class}
A class is defined using the \texttt{class} keyword.

\begin{lstlisting}[language=python]
class Person:
    def __init__(self, name, age):
        self.name = name
        self.age = age

    def greet(self):
        print(f"Hello, my name is {self.name} and I am {self.age} years old.")
\end{lstlisting}
\end{frame}


\begin{frame}{Defining a Class}
\begin{itemize}
\item \texttt{\_\_init\_\_} is the constructor method, called when a new object is created.
\item \texttt{self} refers to the current instance of the class.
\item Attributes like \texttt{name} and \texttt{age} are set on the instance.
\item \texttt{greet} is a method that uses instance attributes.
 \end{itemize}
 \end{frame}



\begin{frame}[fragile]
\frametitle{Creating an Object}
You can create and use an object like this:

\begin{lstlisting}[language=Python]
p = Person("Alice", 30)
p.greet()
\end{lstlisting}

\end{frame}


\begin{frame}[fragile]
\frametitle{Class vs Instance Variables}
\begin{lstlisting}[language=python]
class Dog:
    species = "Canis familiaris"  # Class variable

    def __init__(self, name):
        self.name = name  # Instance variable
\end{lstlisting}
\end{frame}



\begin{frame}[fragile]
\frametitle{Inheritance}
Classes can inherit from other classes:

\begin{lstlisting}[language=python]
class Animal:
    def speak(self):
        print("Makes a sound")

class Dog(Animal):
    def speak(self):
        print("Barks")
\end{lstlisting}
\end{frame}

\begin{frame}[fragile]
\frametitle{Encapsulation}
Encapsulation restricts direct access to some components:
\begin{lstlisting}[language=python]
class BankAccount:
    def __init__(self, balance):
        self.__balance = balance  # Private variable

    def deposit(self, amount):
        self.__balance += amount

    def get_balance(self):
        return self.__balance
\end{lstlisting}
\end{frame}

\begin{frame}{Conclusion}
Python classes offer a powerful way to model real-world entities by combining data and behavior. Key concepts include constructors, instance and class variables, inheritance, and encapsulation.
\end{frame}
