\documentclass[mathserif,20pt,xcolor=table,compress,aspectratio=169]{beamer}

\def\titleskip{2.8cm}
\input{template}

\title[]{\large K-Means Clustering}
%\subtitle[]{}
\author[]{Andreas Mang}
\institute[]{Department of Mathematics, Scientific Computing, Optimization, and Parallel Algorithms Lab, University of Houston}
\date[]{Python Workshop, Department of Mathematics, University of Houston}

\begin{document}

\begin{frame}[plain,label=mytitlepage]
\titlepage
\end{frame}


\begin{frame}{What is K-Means Clustering?}
\begin{itemize}
    \item An unsupervised learning algorithm.
    \item Partitions data into $k$ clusters.
    \item Each point belongs to the cluster with the nearest centroid.
    \item Objective: minimize intra-cluster variance.
\end{itemize}
\end{frame}

\begin{frame}{Algorithm Overview}
\begin{enumerate}
    \item Initialize $k$ centroids randomly.
    \item Assign each point to the nearest centroid.
    \item Update centroids as the mean of assigned points.
    \item Repeat steps 2–3 until convergence.
\end{enumerate}
\end{frame}

\begin{frame}[fragile]{Python Libraries}
\begin{lstlisting}[language=Python]
import numpy as np
import matplotlib.pyplot as plt
from sklearn.cluster import KMeans
from sklearn.datasets import make_blobs
\end{lstlisting}
\end{frame}

\begin{frame}[fragile]{Generating Synthetic Data}
\begin{lstlisting}[language=Python]
X, y_true = make_blobs(n_samples=300,
                       centers=4,
                       cluster_std=0.60,
                       random_state=0)

plt.scatter(X[:, 0], X[:, 1], s=50)
plt.title("Generated Data")
plt.show()
\end{lstlisting}
\end{frame}

\begin{frame}[fragile]{Applying K-Means}
\begin{lstlisting}[language=Python]
kmeans = KMeans(n_clusters=4, random_state=0)
kmeans.fit(X)

y_kmeans = kmeans.predict(X)
centroids = kmeans.cluster_centers_
\end{lstlisting}
\end{frame}

\begin{frame}[fragile]{Visualizing Clusters}
\begin{lstlisting}[language=Python]
plt.scatter(X[:, 0], X[:, 1], c=y_kmeans, cmap='viridis')
plt.scatter(centroids[:, 0], centroids[:, 1],
            c='red', s=200, marker='X')
plt.title("K-Means Clustering Results")
plt.show()
\end{lstlisting}
\end{frame}

\begin{frame}[fragile]{Choosing $k$: The Elbow Method}
\begin{lstlisting}[language=Python]
inertia = []
for k in range(1, 10):
    km = KMeans(n_clusters=k, random_state=0)
    km.fit(X)
    inertia.append(km.inertia_)

plt.plot(range(1, 10), inertia, marker='o')
plt.title("Elbow Method")
plt.xlabel("k")
plt.ylabel("Inertia")
plt.show()
\end{lstlisting}
\end{frame}


\begin{frame}{What is the Elbow Method?}
\begin{itemize}
    \item Determine optimal number of clusters $k$.
    \item Compute K-Means for a range of $k$ values.
    \item Plot the \textbf{inertia} (sum of squared distances to centroids).
    \item Look for an "elbow" in the plot where inertia drops off more slowly.
    \item The $k$ at the elbow point is considered a good choice.
\end{itemize}
\end{frame}

\begin{frame}{Key Concepts}
\begin{itemize}
    \item \textbf{Centroids}: Mean of data points in a cluster.
    \item \textbf{Inertia}: Sum of squared distances from each point to its centroid.
    \item \textbf{Convergence}: When centroids no longer move.
\end{itemize}
\end{frame}

\begin{frame}{Limitations of K-Means}
\begin{itemize}
    \item Requires specifying $k$ in advance.
    \item Sensitive to initial centroid placement.
    \item Assumes spherical clusters of similar size.
\end{itemize}
\end{frame}

\begin{frame}{Conclusion}
\begin{itemize}
    \item K-Means is simple and efficient for clustering.
    \item Best used when clusters are well-separated.
    \item Combine with techniques like the Elbow Method for choosing $k$.
\end{itemize}
\end{frame}

\end{document}

