\documentclass[mathserif,20pt,xcolor=table,compress,aspectratio=169]{beamer}

\def\titleskip{2.8cm}

\input{template}

%\finaltrue
%\showfigstrue
%\showanimtrue

\title[]{\large Bayesian Inverse Problems and Uncertainty Quantification}
\subtitle[]{Math 6397 }
\author[]{Andreas Mang}
\institute[]{SCOPA Group\\Department of Mathematics, University of Houston}
\date[]{\today}

\begin{document}

\begin{frame}[plain,label=mytitlepage]
\titlepage
\end{frame}



\begin{frame}{Contact Information}
\begin{small}
\begin{description}[Instructor]
\setlength\itemsep{0.2em}
\item[Instructor] Prof.~Dr.~Andreas Mang
\item[Email] \href{mailto:andreas@math.uh.edu}{andreas@math.uh.edu}
\item[Website] \href{https://www.math.uh.edu/~andreas}{https://www.math.uh.edu/$\sim$andreas}
\item[Office] PGH 614
\end{description}
\end{small}
\end{frame}



\begin{frame}{Office Hours}
\begin{enumerate}
\setlength\itemsep{0.2em}
\item TuTh 10:00\,am--11:00\,am.
\item By appointment (\href{mailto:andreas@math.uh.edu}{andreas@math.uh.edu}).
\end{enumerate}
\end{frame}



\begin{frame}{Course Website and Material}
\begin{itemize}
\item This is a research-oriented graduate course.
\item Course material, homework assignments, and important announcements will be posted on \textbf{blackboard}. Please visit it on a regular basis.
\item The syllabus can be found on my webpage: \href{https://www.math.uh.edu/~andreas}{https://www.math.uh.edu/$\sim$andreas}
\end{itemize}
\begin{center}
\href{https://www.math.uh.edu/~andreas/teaching/math6397-SP23}{{\color{red}\textit{andreas@math}}/teaching/math6397-SP23}
\end{center}

\end{frame}



\begin{frame}{Some Rules / Suggestions}
\begin{itemize}
\item Read syllabus and check course website.
\item Come to class on time.
\item Attendance is \emph{not} mandatory, but encouraged.
\item Study with students in this course / section.
\item Ask and answer questions (I'm here to make you succeed).
\end{itemize}
\end{frame}



\begin{frame}{Goals}
Connect some research areas:
\begin{itemize}
\item Inverse Problems
\item Monte Carlo sampling
\item High-dimensional probability
\item Machine learning
%\item Gaussian process regression
\item Optimization
\item Computational aspects
\end{itemize}
\end{frame}


\begin{frame}{Content}
\begin{itemize}
\item Intro to IP/regularization theory
\item Intro to probability (Bayesian statistics)
\item (Non)Linear statistical IP
\item MAP estimation
\item MCMC sampling and beyond
\item $\ldots$
\end{itemize}
\end{frame}


\begin{frame}{Required Work}
\begin{itemize}
\item $n$ homework assignments (mix of paper and pencil and computations)
\item if $n>2$: You can replace one assignment by a final project ideally related to your research.
\end{itemize}
\end{frame}

\begin{frame}{Required Work}
\begin{itemize}
\item Code can be written in your preferred language/programming environment (Python; Matlab; Julia; Mathematica; $\ldots$)
\item Matlab resources available online ${(\ast)}$
\end{itemize}
${(\ast)}$: \small\url{https://github.com/andreasmang/machine}
\end{frame}


\end{document}
